How has the Brazilian Amazon been constructed as a problem:Presidential speeches and transnational politics since 1985

### 2022-09-20 Draft_September_2022/Full_draft_20220901.rmd
#added references

---
title: How has the Brazilian Amazon been constructed as a problem:Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "12 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about and act towards the environment transnationally, but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses across time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches inside and outside Brazil. While constructing the Amazon as a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest away presidents are from the Amazon itself, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

@@

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices, are a crucial factor in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different settings, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

@@

@hecht1990 [p. 1] write that "what imbues the case of the Amazon with such passion is the symbolic content of the dreams it ignites". In the process of securing Amazonian borders since the 18th century, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht1990, p. 8], and when the dust settled and the scramble was over, over half of the Amazon emerged Brazilian. While Brazilian military diplomacy was very successful, the process did not come without its traumas as the negotiations with Bolivia in 1902 to secure the territory of the state of Acre [@hecht1990]. These traumas were referred to, and offered context, to Amazonian discourses and policies during the military dictatorship of 1964 aimed at protecting Brazil's sovereignty over the Amazon. As we move from a world where non-state actors gain importance in environmental governance [@silva-muller2022; @andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from state actors to a wide set of actors. Claims of attempts to 'internationalize' the Brazilian Amazon, for instance, are targeted at foreign actors as well as domestic non-state actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem:Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: 20 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about and act towards the environment transnationally, but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses across time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches inside and outside Brazil. While constructing the Amazon as a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest away presidents are from the Amazon itself, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

@@

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015; @rajao2020 ], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices, are a crucial factor in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different settings, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

@@

@hecht1990 [p. 1] write that "what imbues the case of the Amazon with such passion is the symbolic content of the dreams it ignites". In the process of securing Amazonian borders since the 18th century, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht1990, p. 8], and when the dust settled and the scramble was over, over half of the Amazon emerged Brazilian. While Brazilian military diplomacy was very successful, the process did not come without its traumas as the negotiations with Bolivia in 1902 to secure the territory of the state of Acre [@hecht1990]. These traumas were referred to, and offered context, to Amazonian discourses and policies during the military dictatorship of 1964 at earlier regimes aimed at protecting Brazil's sovereignty over the Amazon [@hecht2020]. As we move from a world where non-state actors gain importance in environmental governance [@silva-muller2022; @andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from state actors to a wide set of actors. Claims of attempts to 'internationalize' the Brazilian Amazon, for instance, are targeted at foreign actors as well as domestic non-state actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.
